---
title: "Bonanza Creek Experimental Forest Snowshoe Hares"
author: "German Silva"
date: "11/8/2021"
output: html_document
---

# Introduction


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Attach proper packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
library(calecopal)
```

# Juvenile Snowshoe Hare Data Exploration

## Juvenile Snowshoe Hare Trappings Over Time
 <!-- - Open up the data using `read_csv()`
      - Create Stats
      - Make a graph
      - Obtain Annual Counts
      -->
 
```{r, fig.cap= "**Figure 1** Bar graph showing the amount of juvenile snowshoe hares trapped in Bonanza Creek Experimental Forest over time between 1999 and 2012. The largest number of trappings occured in 1999."}
snowshoe <- read_csv(here("data", "bonanza_hares.csv"))

snowshoe_traps <- snowshoe %>% 
  mutate(date = mdy(date),
         year = year(date)) %>% 
  filter(age == "j")

ggplot(snowshoe_traps, aes(x= year)) +
  geom_bar(fill = "#B4AA98")+
  scale_x_continuous(breaks = 1999:2012) +
  labs(x = "Year of Trapping",
       y = "Number of Snowshoe Hares Traped") +
  ggtitle("Juvenile Snowshoe Hares Trapped Between 1998-2012 (SILVA)")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_rect("#D7DCE4")) 

snowshoe_traps_stats <- snowshoe_traps %>% 
  group_by(year) %>% 
  summarize(count = n())
```
 
**Takeaways**:
The number of juvenile snowshoe hares trapped ranged from `r max(as.vector(snowshoe_traps_stats$count))` in 1999 to `r min(as.vector(snowshoe_traps_stats$count))` in 2010. The average number of trapped juveniles is `r mean(as.vector(snowshoe_traps_stats$count))`, with a median of `r median(as.vector(snowshoe_traps_stats$count))`. There is a general trend of a decreasing number of juveniles being trapped since 1999. To standardize the number of trappings (due to effort), trapping campaigns should be a set length of 5 days with a standard number of traps (n=30) with traps being checked twice daily and hares being measured whenever encountered.

## Sex Differences in Juvenile Snowshoe Hare Weight

```{r}
ggplot(snowshoe_traps, aes(x= sex, y= weight, color = sex)) +
  geom_jitter()+
  facet_wrap(~grid)
```

